### 2013-03-28 doc/sosp13/intro.tex
move stuff out

In this paper we introduce \Grappa, a software runtime system that allows a commodity cluster of x86-based nodes connected via an InfiniBand network to be programmed as if it were a single, large, shared-memory NUMA (non-uniform memory access) machine with scalable performance for irregular applications. \Grappa leverages as much freely available and commodity infrastructure as possible. We use unmodified Linux for the operating system and an off-the-shelf user-mode InfiniBand device driver stack~\cite{OFED}. MPI is used for process setup and tear down. GASNet~\cite{gasnet} is used as the underlying mechanism for remote memory reads and writes using active message invocations. \Grappa adds three main software components: (1) a \emph{lightweight tasking\/} layer that supports a context switch in as little as 38ns and distributed global load balancing; (2) a \emph{distributed shared memory\/} layer that supports normal access operations such as \emph{read\/} and \emph{write\/} as well as synchronizing operations such as \emph{fetch-and-add\/}~\cite{fetchandadd}; and (3) a \emph{communication\/} layer that combines short messages to achieve peak bandwidth on commodity networks. As we will show later, \Grappa can tolerate latencies way beyond that of the network. Therefore, \Grappa can afford to \emph{trade latency for throughput}. By {\em increasing\/} latency in key components of the system we are able to: increase effective random access memory bandwidth by delaying and aggregating messages; increase synchronization rate by delegating atomic operations to gatekeeper cores, even when referencing node-local global data; and improve load balance by tolerating the delays incurred when work-stealing.

>>>>

The two points above form the cornerstone of \Grappa, a software runtime
system that allows a commodity cluster of x86-based nodes connected via an
InfiniBand network to be programmed as if it were a single, large,
shared-memory NUMA (non-uniform memory access) machine with scalable
performance for irregular applications. \Grappa tolerates communication
latencies with lightweight user-level tasking layer that supports a context
switch in as little as 38ns and can sustain a large number of active contexts.
It bridges the commodity network bandwidth gap with a communication layer that
combines short messages into larger ones.

% In this paper we introduce \Grappa, a software runtime system that allows a
% commodity cluster of x86-based nodes connected via an InfiniBand network to be
% programmed as if it were a single, large, shared-memory NUMA (non-uniform
% memory access) machine with scalable performance for irregular applications.
% \Grappa tolerates communication latencies with lightweight user-level tasking layer that supports a context switch in as little as 38ns and can sustain a large number of active contexts. It bridges the commodity network bandwidth gap with a communication layer that combines short messages into larger ones. 
% 
% , and it addresses the a \emph{communication\/}
% layer that combines short messages to achieve peak bandwidth on commodity
% networks
% 
% 
% distributed global load balancing
% 
% 
% \Grappa leverages as much freely available and commodity infrastructure as
% possible. We use unmodified Linux for the operating system and an
% off-the-shelf user-mode InfiniBand device driver stack~\cite{OFED}. MPI is
% used for process setup and tear down. GASNet~\cite{gasnet} is used as the
% underlying mechanism for remote memory reads and writes using active message
% invocations.
% 
% \Grappa adds three main software components: (1) a \emph{lightweight
% tasking\/} layer that supports a context switch in as little as 38ns and
% distributed global load balancing; (2) a \emph{distributed shared memory\/}
% layer that supports normal access operations such as \emph{read\/} and
% \emph{write\/} as well as synchronizing operations such as
% \emph{fetch-and-add\/}~\cite{fetchandadd}; and (3) a \emph{communication\/}
% layer that combines short messages to achieve peak bandwidth on commodity
% networks. As we will show later, \Grappa can tolerate latencies way beyond
% that of the network. Therefore, \Grappa can afford to \emph{trade latency for
% throughput}. By {\em increasing\/} latency in key components of the system we
% are able to: increase effective random access memory bandwidth by delaying and
% aggregating messages; increase synchronization rate by delegating atomic
% operations to gatekeeper cores, even when referencing node-local global data;
% and improve load balance by tolerating the delays incurred when work-stealing.
